Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1)

### 2014-04-25 WSSSPE1_report/WSSSPE1_summary_report.tex
Abstract: refine.

\begin{abstract} %\note{Lead: Sou-Cheng T. Choi}
% Anne: Added a lead-in motivator to address the submission as a Experience paper
% and tried to improve writing... ( I am assumeing no abstract word count limit...)
Challenges related to the development, deployment, and maintenance of 
reusable software for science is becoming an increasing concern. 
There are now many scientists whose research increasingly depends on 
the quality and availability of software they can use and build on. 
At the same time, software specialists have more than enough to do 
without having to "reinvent the wheel". To highlight some of these 
issues and share experiences, The First Workshop on Sustainable 
Software for Science: Practice and Experiences (WSSSPE1) was held 
in November 2013 at SC13.
The workshop featured keynote presentations and based its discussion 
them, as well as on a large number (54) of solicited extended abstracts 
that were grouped into three themes presented via panels that encouraged 
audience participation. A set of collaborative notes was taken during the workshop.

This document 
% article? 
summarizes 
the issues and shared experiences discussed atthe workshop, including
cross-cutting issues and use cases. It joins a nascent
literature seeking to understand what drives software work in science,
and how it is impacted by the reward systems of science. 
These incentives can determine the extent to which
developers are motivated to build software for the long-term, for the
use of others, and whether to work collaboratively or separately.

Unique perspectives were captured about issues such as comprehensive
documentation, development and deployment practices, software
licenses and career paths for developers. Attribution systems that capture
evidence of software contribution and impact were also discussed. These include
mechanisms such as Digital Object Identifiers, publication
of ``software papers'' and the use of online systems, for example source code repositories
like GitHub. For truly sustainable software, there should be no
endpoint, as the software products continue to be used beyond a single
institution, grant, and developer or development team.

>>>>

\begin{abstract} %\note{Lead: Sou-Cheng T. Choi}
% Anne: Added a lead-in motivator to address the submission as a Experience paper
% and tried to improve writing... ( I am assumeing no abstract word count limit...)
Challenges related to the development, deployment, and maintenance of 
reusable software for science are becoming a growing concern. 
There are now many scientists whose research increasingly depends on 
the quality and availability of software they can use and build on. 
At the same time, software specialists have more than enough to do 
without having to ``reinvent the wheel.'' To highlight some of these 
issues and share experiences, The First Workshop on Sustainable 
Software for Science: Practice and Experiences (WSSSPE1) was held 
in November 2013 at SC13.
The workshop featured keynote presentations and based its discussion 
on them, as well as on a large number (54) of solicited extended abstracts 
that were grouped into three themes presented via panels that encouraged 
audience participation. A set of collaborative notes was taken during the workshop.

This document summarizes 
the issues and shared experiences discussed at the workshop, including
cross-cutting issues and use cases. It joins a nascent
literature seeking to understand what drives software work in science,
and how it is impacted by the reward systems of science. 
These incentives can determine the extent to which
developers are motivated to build software for the long-term, for the
use of others, and whether to work collaboratively or separately. We 
also elabroate on community building, leadership, and dynamics leading to 
successful scientific software.

Unique perspectives were captured about issues such as comprehensive
documentation, development and deployment practices, software
licenses and career paths for developers. Attribution systems that accounts for 
evidence of software contribution and impact were also discussed. These include
mechanisms such as Digital Object Identifiers, publication
of ``software papers'', and the use of online systems, for example source code repositories
like GitHub. For truly sustainable software, there should be no
endpoint, as the software products continue to be used beyond a single
institution, grant, and developer or development team.
